---
title: "Datatables"
output:
  slidy_presentation:
    theme: default
css: "www/teachingnotes.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	eval = FALSE,
	message = FALSE,
	warning = FALSE,
	include = TRUE
)
```

## Displaying data with datatables

Datatable JS is one of the most popular libraries for creating interactive, responsively designed tables on the web.

Often it is crucial to display your actual raw data as a table, and to make it downloadable as a .csv/.xlsx easily. Here are two examples of where I've found this to be the case:

- [Oxford Roman Economy Project - Shipwrecks Database](https://livedataoxford.shinyapps.io/1499182601-r5pqgdongyjr1qo/)
- [HBV Epitopes Database](http://www.expmedndm.ox.ac.uk/hepitopes)

## Using datatable to display data (I)

Let's load up the datatable-and-the-journey.RProj project from RSS_handouts folder and visualise the journeys dataset.

To create a table, we simply use:

```{r}
library("DT")
datatable(journeys)
```

Let's create a table with the following specifications:

- Show only the start and end locations, date, and number of letters sent
- Remove periods from the column names
- Hide the row names

## Using datatable to display data (II)

Datatables include by default the following elements:

- Show X entries
- The table
- Pagination info
- Pagination

We can change which elements appear (and it what order) using the `dom` argument of options:

```{r}
datatable(journeys,
          options = list(
            dom = "iftlp"
          ))
```

## Customising column appearance

There are a number of utility functions for customising individual columns, these are used very similarly to one another:

```{r}
journeys %>%
  select(start.location,
         end.location,
         date,
         number.of.letters) %>%
  datatable(filter = "top",
            options = list(dom = "iftlp")) %>%
  formatCurrency("number.of.letters", currency = "$", mark = ",") %>%
  formatStyle("date", fontWeight = "bold")
```

## Learning more about DT

I thoroughly recommend this website maintained by the developers of the DT htmlwidget (RStudio!): [rstudio.github.io/DT](https://rstudio.github.io/DT)

