---
title: "Highcharter"
output:
  slidy_presentation:
    theme: default
css: "www/teachingnotes.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	eval = FALSE,
	message = FALSE,
	warning = FALSE,
	include = TRUE
)
```

## Highcharter and Highcharts JS

Highcharter utilises the Highcharts JS library to create beautifully, very minimal charts. It's my personal favourite option for creating simple charts.

However, it comes with a huge disclaimer:

```{r, eval=TRUE, message=TRUE, warning=TRUE}
library("highcharter")
```

## Highcharter licensing

For more information about licensing, see https://shop.highsoft.com/highcharts.

It should be noted that the Highcharts company really like the `highcharter` library, they even provide a discount on Single Developer licenses to users of the library: http://announcements.highcharts.com/foss/.

## Highcharter compared to plotly and leaflet

`highcharter` is different from `plotly` and `leaflet` in that it doesn't use the ~ syntax to access columns.

There are two very different ways to create 

<div class="row">
<div class="col-md-6">
**Charts from scratch**

- Instantiate a highcharter object with `highchart`
- Add series with `hc_add_series` 
- Ensure column names within `hc_add_series` are consistent with the Highchart JS library's requirements.

</div>

<div class="col-md-6">
**Charts with magic**

- Provide a tibble to the `hchart` function
- Specify the chart aesthetics using `hcaes`
- Provide column names as "naked variables"

</div>
</div>

<br>
We're going to use the magic of `hchart` in this course, but I'll refer you to documentation for the more explicit process of building `highcharter` charts later.

## Highcharter and Gapminder (I)

Let's create the same chart we created previously, but in highcharter:

```{r}
library("tidyverse")
library("gapminder")
library("highcharter")
continent_mean_pop_growth <- gapminder %>%
  group_by(continent, year) %>%
  summarise(mean.pop = mean(pop))
continent_mean_pop_growth %>%
  hchart(
    type = "line"
  )
```

## Highcharter and the journeys dataset (I)

Let's load up the highcharter-and-the-journey.RProj project from RSS_handouts folder, there's a file here that processes our data ready to be visualised using a bar chart.

Now let's create this chart using highcharter:

```{r, echo=FALSE, eval=TRUE}
library("tidyverse")
library("forcats")
library("highcharter")

journeys <- read_csv("data/detailed-journeys.csv")

end_location_frequencies <- journeys %>%
  group_by(end.city, end.country) %>%
  count() %>%
  arrange(desc(n)) %>%
  ungroup() %>%
  mutate(end.city = fct_reorder(end.city, n))

end_location_frequencies %>%
  hchart(
    type = "bar",
    hcaes(
      x = end.city,
      y = n
    )
  )
```

## Highcharter and the journeys dataset (II)

In `plotly` we coloured each city according to the country it belonged to, we could naively achieve this in `highcharter` as follows:

```{r, eval=TRUE}
end_location_frequencies %>%
  hchart(
    type = "bar",
    hcaes(
      x = end.city,
      y = n,
      group = end.country
    )
  )
```

Unfortunately, `highcharter` is far too clever for its own good and splits the data into two groups.

The simplest method to fix this is as follows:

```{r, eval = TRUE}
end_location_frequencies %>%
  mutate(position = row_number() - 1) %>%
  hchart(
    type = "bar",
    hcaes(
      x = position,
      y = n,
      group = end.country
    )
  ) %>%
  hc_xAxis(categories = end_location_frequencies$end.city)
```

## Highcharter and the journeys dataset (III)

The solution provided before isn't perfect, as noted by the package author here: [https://github.com/jbkunst/highcharter/issues/230#issuecomment-268715227](https://github.com/jbkunst/highcharter/issues/230#issuecomment-268715227).

This highlights nicely a lesson you must learn with htmlwidgets; create a priority list for the features of a chart, and accept there will be trade-offs to chasing particular functionalities.

## Learning more about highcharter

There's an excellent website managed by the package developer here: [http://jkunst.com/highcharter/](http://jkunst.com/highcharter/).

The package is also deliberately designed to be as close to the underlying Highcharts JS library as possible, I therefore recommend you consult the API documentation for advanced features: [http://api.highcharts.com/highcharts](http://api.highcharts.com/highcharts)


