---
title: "Shiny Advice"
output: slidy_presentation
runtime: shiny
css: "www/teachingnotes.css"
---


# Embedding "Split File" Shiny Apps into Presentations

There are two methods of embedding split file apps into RMarkdown presentations:

>- `shinyAppDir`: requires the entire RMarkdown document to be in the shiny runtime, i.e. shinyapps.io active hours are consumed while anywhere in the presentation

>- iframes: allows a shinyapps.io hosted Shiny app to be embedded into a single (or many) slides without the rest of the app requiring Shiny active hours to display

# Designing Shiny apps

Shiny apps that you wish to deploy to shinyapps.io (or your own server) should be split into two files:

- ui.R
- server.R

Data for your shiny apps that *needs* to be included with the app should be stored in a sub-directory called "data".

This means your data files can easily be accessed with the path: "data/my_file.csv"

# Images and more in Shiny apps

Shiny makes it easy to add images (as well as CSS and JavaScript) to your apps, provided that these files are stored within a sub-directory called "www"

In the exercise folder there is a directory called "shinyAdvice_app-structure" which demonstrates this layout.

# Split code into manageable files

In large Shiny apps, particurly `navbarPage` layouts, it makes sense to split different sections of the `server.R` code into separate files. These can then be called using:

```{r, eval=FALSE}
## load once, not dependent on input actions
source("data-processing.R", local = TRUE)

## reactive section of server
shinyServer(
  function(input, output, session){
    
    source("maps.R", local = TRUE)$value
    
  }
)
```

# Write functions instead of reusing code

If you're doing the same thing multiple times, write a function.

Functionalising code provides the following advantages:

>- Makes your code easier to read
>- Make your code easier to debug (comment out function calls)
>- Makes your code easier to update

<div class="incremental">
Here's an example:

```{r, eval=FALSE}
output$capital_map <- renderLeaflet(
  leaflet(data = capitals()) %>%
    addProviderTiles("Stoner.Tone") %>%
    addCircleMarkers(popup = ~name, weight = ~size) %>%
    addLegend(position = "topright") %>%
    addMarkers(lng = country.centre.lng, lat = country.centre.lat)
)

output$famous_attractions_map <- renderLeaflet(
  leaflet(data = famous_attractions()) %>%
    addProviderTiles("Stoner.Tone") %>%
    addCircleMarkers(popup = ~name, weight = ~size) %>%
    addLegend(position = "topright") %>%
    addMarkers(lng = country.centre.lng, lat = country.centre.lat)
)
```

```{r, eval=FALSE}
custom_map <- function(locations.data = NA){
  leaflet(locations.data) %>%
    addProviderTiles("Stoner.Tone") %>%
    addCircleMarkers(popup = ~name, weight = ~size) %>%
    addLegend(position = "topright") %>%
    addMarkers(lng = country.centre.lng, lat = country.centre.lat)
}

output$capital_map <- renderLeaflet(
  custom_map(capitals())
)

output$famous_attractions_map <- renderLeaflet(
  custom_map(famous_attractions())
)
```
</div>

# Some HTML you have to know

Unfortunately, it's not quite true you don't need to know any HTML to use Shiny. But there are only a few things you need to know.

# No duplicate input/outputs

All `*Input` functions get converted into code like this: 

```{HTML}
<div class="form-group shiny-input-container">
  <label class="control-label" for="thisVar">label</label>
  <div>
    <select id="thisVar"><option value="1" selected>1</option>
<option value="2">2</option>
<option value="3">3</option></select>
    <script type="application/json" data-for="thisVar" data-nonempty="">{}</script>
  </div>
</div>
```

There can only be **one** instance of any \<select id="thisVar">\. You cannot use the same input variable more than once.

All `*Output` functions get converted into code like this:

```{HTML}
<div id="capital_city_map">

</div>
```

There can only be **one** instance of any \<div id="thisVar">\. You cannot use the same output object more than once.

# No duplicate input/outputs

If you do need to reuse output multiple times, then the simplest solution is as follows:

```{r, eval=FALSE}
output$hist_1 <- output$hist_2 <- output$hist_3 <- renderPlot(
  hist(rnorm(n = 1000))
)
```

# New lines

There are two different ways to put content across multiple lines in HTML:

- \<p>This is the paragraph tag and puts content within it on a new line, though it should only really be used for separating paragraphs\</p>
- This is the line break character, \<br>\, it forces a new line

These are very useful within htmlwidgets

```{r}
library(readr)
library(leaflet)
capital_cities <- read_csv("data/capitals_with_locations.csv")
capital_cities %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = ~paste("Capital:", Capital,
                            "<br>", 
                            "Country:",Country)
             )
```

# Links

Links are added using the \<a>\ tag. Which has a convenient wrapper in Shiny:

```{r, eval=FALSE}
a(href = "http://google.com", "Link to Google")
```

# Getting Help

[Stackoverflow](http://stackoverflow.com/) is the Q&A community for programming and scripting, this is a good place to start when trying to solve a problem. Note that the community may see abrasive to new users, the following advice is useful to prevent feeling like your fingers have been snapped off:

- Search for a solution before asking a question
- Provide a [reproducible example of your problem](http://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-example)

There are also additional training courses on R run by the University, in addition to training courses on [Lynda.com](https://shib.lynda.com/Shibboleth.sso/InCommon?providerId=https://registry.shibboleth.ox.ac.uk/idp).

# Getting Help

Research Support Services are currently launching a visualisation support service, called the Interactive Data Network.

Shiny is going to be a fully supported platform for this service, with tutorials and consultancy provided to researchers on a recharge basis. Please contact martin.hadley@it.ox.ac.uk at present for further information.

Tutorial collections you may find useful are available here:

- http://ox-it.github.io/OxfordIDN_htmlwidgets/
- https://ox-it.github.io/OxfordIDN_Shiny-App-Templates/

# Example of Shiny Apps

The following resources provide great examples of what you can do with Shiny:

- shiny.rstudio.com/gallery/
- http://www.showmeshiny.com/
- daattali.com/shiny/




