---
title: "RSS: Interactive Visualisations with R"
subtitle: "Day One: Tidyverse, htmlwidgets and RMarkdown"
output:
  slidy_presentation:
    theme: default
css: "www/teachingnotes.css"
---

```{r, include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

# Timetable for the day

Today's timetable is roughly as follows:

```{r}
timetable <- list(
  "09:30 - 10:30" = "Up and Running with R and RStudio",
  "10:30 - 11:00" = "Setting ourselves up for reproducible data visualisations",
  "11:00 - 11:20" = "COFFEE",
  "11:20 - 12:00" = "Wrangling data and Exploratory Data Analysis in RStudio with the Tidyverse",
  "12:00 - 13:00" = "Interactive maps with htmlwidgets",
  "13:00 - 13:45" = "LUNCH",
  "13:45 - 14:30" = "Interactive charts and network diagrams with htmlwidgets",
  "14:30 - 15:00" = "Interactive reports and presentations with RMarkdown",
  "15:00 - 15:20" = "COFFEE",
  "15:20 - 16:00" = "Inserting htmlwidgets into RMarkdown",
  "16:00 - 17:00" = "RMarkdown to the web (and back)"
)
library("tidyverse")
timetable <- tibble(
  times = names(timetable),
  topic = as.character(timetable)
)
knitr::kable(timetable)
```

# Notes on course expectations

This course is designed to introduce you to the basics of creating interactive data visualisations using R; with two very different technologies:

No prior knowledge of R is expected, the necessary "programming patterns" and workflows will be introduced through the course.

Please note that this course will **NOT** provide a sufficient overview of the R language to start implementing complex analyses or how to write efficient programs with R.

>* This course shows you how to make Shiny stuff.
