---
title: "htmlwidgets"
author: "Martin John Hadley"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

# htmlwidgets: JavaScript

The interactive web is driven by JavaScript*.

The majority of interactive elements that you use on websites are written in JavaScript - from interactive maps to auto-completing pop-up menus. 

.pull-left[

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library("tidyverse")
library("gapminder")
library("highcharter")
library("lubridate")

continent_mean_pop_growth <- gapminder %>%
  group_by(continent, year) %>%
  summarise(mean.pop = mean(pop)) %>%
  mutate(year = dmy(paste0("31-12-", year)))

highchart(type = "stock") %>%
  hc_add_series(
    data = continent_mean_pop_growth,
    type = "line",
    hcaes(
      x = year,
      y = mean.pop,
      group = continent
    ),
    marker = list(
      enabled = TRUE
    )
  ) %>%
  hc_legend(enabled = TRUE) %>%
  hc_rangeSelector(enabled = FALSE) %>%
  hc_scrollbar(enabled = FALSE) %>%
  hc_navigator(enabled = FALSE) %>%
  hc_title(text = "Continental Populations<br> (1952 - 2007)") %>%
  hc_size(width = "100%", height = "300px")

```
]

.pull-right[
```{r, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library("tidyverse")
library("leaflet")
library("sf")
library("gapminder")

geojson_worldmap <- st_read("https://raw.githubusercontent.com/johan/world.geo.json/master/countries.geo.json",
                            quiet = TRUE)
geojson_worldmap <- invisible(geojson_worldmap %>%
  left_join(gapminder %>%
               filter(year == max(year)) , by = c("name" = "country")) %>%
  filter(name != "Antarctica"))
palette_pop <- colorNumeric("YlOrBr" , geojson_worldmap$pop, na.color = "#c0c1c4")
geojson_worldmap %>%
  leaflet(width = "300px",
          height = "200px") %>%
  addPolygons(fillColor = ~palette_pop(pop),
              fillOpacity = 0.8,
              color = "#000",
              weight = 1,
              popup = ~name)
```
]


---

# JavaScript Libraries

As with R, there are hundreds of different JavaScript libraries designed for various visualisation tasks.

Examples of these include:

- D3

- plotly

- highcharts

- vis.js

- Google Charts

If you get into conversation with JavaScript developers you might get into semantic arguments about libraries/frameworks. 

We really don't care about these here.

---

# htmlwidgets: The Framework

`highcharts` is a framework for building bindings between JavaScript libraries and the R language.

R developers use the `highcharts` package to scaffold together R packages that allow R users [useRs] to build interactive things directly from R code.

---

# htmlwidgets: The Packages

These R packages work like any other package, and there are many examples available on CRAN.

Typically, an R package binds to just **one** JavaScript library.

- `plotly` allows us to create plotly charts

- `highcharter` allows us to create highcharts charts

- `visNetwork` and `timevis` both bind to the visjs library but are used to create very different charts; network visualisations and Gantt charts, respectively.

---

# htmlwidgets.org

The spiritual home of `htmlwidgets` is [htmlwidgets.org](http://htmlwidgets.org) and is maintained by the folks at RStudio.

- [htmlwidgets.org](http://htmlwidgets.org) introduces the basics of building your own `htmlwidget` packages<sup>1</sup>

- [htmlwidgets.org](http://htmlwidgets.org) has a showcase of the 13 most high-profile `htmlwidget` packages (many of which have been developed by RStudio themselves)

- [htmlwidgets.org](http://htmlwidgets.org) has a gallery of many other `htmlwidget` packages.

.footnote[
[1] This is faaaar beyond the scope of today's course
]

---

# htmlwidgets sanity check

Every single `htmlwidget` package is completely unique in the way that data and chart options are specified.

Why?

- Underlying JavaScript libraries/frameworks are often completely different to one another.

- Developers have a choice between an easy to use `htmlwidget` and a package that allows you to do *everything* you could do with the native JS library.

- Developers have a choice in how `tibble` columns are specified to the chart; `"columns_as_strings"`, `naked_columns`, `~formula_columns`.
---

# Which libraries will we cover?

We're going to introduce 6 very different `htmlwidgets` which fit (fairly) neatly into the following categories:

- Maps

  - leaflet
  - mapview
  
- Charts

  - highcharter
  - plotly
  
- Networks

  - visNetwork
  
- Tables

  - DT
