---
title: "Interactive Maps with htmlwidgets"
author: "Martin John Hadley"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

# Interactive maps with htmlwidgets

We're going to introduce three libraries for creating interactive maps:

- leaflet: a powerful library for creating highly customised presentation quality charts.

- mapview: geared towards exploratory data analysis of geographical data.

- mapedit: an extremely powerful tool for modifying geographical data.

<br>

Before we introduce these libraries, we must first discuss how to handle GIS data in R.

---

# rgdal, sp and sf

If you've done GIS with R over the last few years, it's more than likely you've used a combination of the rgdal and sp packages.

It's highly recommended that from now on you use the sf package, and if possible backport your code to sf.

<hr>

You will likely come across StackOverflow questions etc that continue to use rgdal and sp, it's definitely worth the effort to work exclusively in sf. 

Plus... it's an easy way for you to contribute to the R Community with sp dependent solutions to question

---

# Why use sf?

sf provides a consistent approach to working with GIS data in R, allowing us to do the following very easily:

- import GIS data

- wrangle GIS data using the tidyverse

- perform geometric computations

- visualise with ggplot, leaflet, mapview and mapedit

Initial and future development of the package is partly funded and supported by the R Consortium.

Read more about the package here [https://r-spatial.github.io/sf/](https://r-spatial.github.io/sf/)

---

# Working with shapefiles in sf

Shapefiles are GIS datasets defining geographic regions, i.e. country borders.

In the [martinjhnhadley/collated-rstats-training-materials](https://goo.gl/2LwAWh) repo [[https://goo.gl/2LwAWh](https://goo.gl/2LwAWh)] there's an RStudio project that helps us introduce how to work with shapefiles with sf:

1. Open the sf-examples.RProj project

1. Open the import-sf.R file

1. Run these lines:

```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("sf")

world_shapefiles <- read_sf("data/world-shape-files/")
```

---

# Augmented tibbles

```{r}
class(world_shapefiles)
```


---

# Pretty printing

```{r}
world_shapefiles
```

---

# Wrangling sf objects

```{r}
world_shapefiles %>%
  filter(CONTINENT == "Europe")
```

---

# Tip: Inconsistent column headings

It's really frustrating working with datasets with inconsistent column headings, I recommend always doing the following:

```{r}
world_shapefiles <- world_shapefiles %>%
  setNames(. , tolower(colnames(.)))
colnames(world_shapefiles)
```

---

# mapedit

mapedit is a powerful tool for modifying and augmenting sf objects:

```{r, eval = FALSE}
library("tidyverse")
library("statesRcontiguous")
library("mapedit")

selected_states <- shp_all_us_states %>%
  selectFeatures()
```

Sadly, `mapedit` elements don't play nicely in this presentation mode.

---

# mapview

`mapview` provides all of the visual components of the `mapedit` package:

```{r, eval=FALSE}
library("tidyverse")
library("statesRcontiguous")
library("mapview")

shp_all_us_states %>%
  mapview()
```

Sadly, `mapview` elements don't play nicely in this presentation mode.

---

# Leaflet

Leaflet allows us to make beautiful maps, it's way more fully featured than `mapview`.

```{r, echo=FALSE}
library("leaflet")
library("widgetframe")
lf_quakes <- quakes %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  leaflet(width = "100%",
          height = "300px") %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>%
  addCircleMarkers(clusterOptions = markerClusterOptions(),
                   popup = ~paste("Magnitude: ", mag))
lf_quakes
# frameWidget(lf_quakes)
```

Let's deep dive into the library.

---

# quakes

There's a really useful example dataset in base R called `quakes`

```{r}
quakes %>%
  head()
```

Let's convert this into an `sf` object.

---

`st_as_sf` needs the following information

- `coords`: these must be given in longitude, latitude order.
- `crs`: the coordinate references system, it's sensible to always use 4326 as a default.

```{r}
quakes %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  head()
```

---

# quakes map

Just as with `ggplot2`, we need to create a leaflet object to which we add layers:

```{r}
quakes %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  leaflet()
```

---

`addProviderTiles` changes the base map ( see [rstudio.github.io/leaflet/basemaps.html](http://rstudio.github.io/leaflet/basemaps.html))

```{r}
quakes %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  leaflet() %>%
  addProviderTiles(providers$Esri.OceanBasemap)
```

---

`addCircleMarkers` adds circles at each data point:

```{r}
quakes %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  leaflet() %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>%
  addCircleMarkers()
```

---

Use `~column.name` to use columns in the visualisation:

```{r}
quakes %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  leaflet() %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>%
  addCircleMarkers(label = ~as.character(mag))
```

---

`markerClusterOptions` allows us to cluster markers:

```{r}
quakes %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326) %>%
  leaflet(width = "100%",
          height = "300px") %>%
  addProviderTiles(providers$Esri.OceanBasemap) %>%
  addCircleMarkers(label = ~as.character(mag),
                   clusterOptions = markerClusterOptions())
```

---

# Publishing htmlwidgets with RPubs

[RPubs.com](rpubs.com) is a free and open platform for publishing and sharing htmlwidgets (and RMarkdown documents).

Its's incredibly simple to publish to RPubs - simply click on the publish button in the htmlwidget viewer!

> **Note that while documents hosted on RPubs are made public, the underlying code/.rmd files cannot be downloaded by others.**

---

# Exercise 1

Create a new project which achieves the following:

1. Reproducibly cleans the raw data from [martinjhnhadley/collated-rstats-training-materials/german-migrants](https://goo.gl/2LwAWh)

1. Import the cleaned data

1. Visualises all the unique locations with `leaflet`

## Bonus Objectives:

- When the cursor hovers over a point, show the name of the city **AND** the country.

- Create a version of the map where send and receive locations are coloured differently.  

---

# Exercise 2

Create a new project which achieves the following:

1. Use the `statesRcontiguous` library to obtain only the **contiguous** USA

1. Visualise this shapefile with `leaflet`

1. Refer to [http://rstudio.github.io/leaflet/legends.html](http://rstudio.github.io/leaflet/legends.html) to colour the states according to the region that they belong to.


