---
title: "Interactive Charts with highcharter"
author: "Martin John Hadley"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library("tidyverse")
library("fivethirtyeight")
tarantino_swears <- tarantino %>%
  filter(complete.cases(.)) %>%
  count(word, movie) %>%
  group_by(word) %>%
  mutate(total.n = sum(n)) %>%
  filter(total.n > 1) 
```

# Highcharter

Highcharter uses the highcharts library to create extremely professional charts.

However, it's extremely important to remember that this library cannot be used freely for commercial purposes:

```{r, eval=FALSE}
library("highcharter")
```

```{r, echo=FALSE}
detach("package:highcharter", unload=TRUE)
library("highcharter")
```

---

# hchart and hcaes

The highcharter library has been designed very much like `gglot2`:

- `hchart` creates a chart
- `hcaes` species the aesthetic mappings, using naked_column_names

```{r, eval=FALSE}
tarantino_swears %>%
  hchart(type = "bar",
         hcaes(
           x = ,
           y = ,
           group = 
         ))
```

---

Highcharter orders the categories as they appear (uniquely) in the dataset:

```{r, echo=FALSE}
tarantino_swears %>%
  hchart(type = "bar",
         hcaes(
           x = word,
           y = n,
           group = movie
         ))
```

---

# highcharter ordering

Fixing the ordering of our bars is slightly complicated, the problem stems from there not being *values* for all combinations of words and movies.

We can fix this with a bit of jerry rigging using `tidyr` and `dplyr`

---

```{r}
tarantino %>%
  filter(complete.cases(.)) %>%
  count(word, movie) %>%
  group_by(word) %>%
  spread(movie, n, fill = 0)
```

---

```{r}
tarantino %>%
  filter(complete.cases(.)) %>%
  count(word, movie) %>%
  group_by(word) %>%
  spread(movie, n, fill = 0) %>%
  gather(movie, count, 2:8)
```

---

```{r}
tarantino_swears <- tarantino %>%
  filter(complete.cases(.)) %>%
  count(word, movie) %>%
  group_by(word) %>%
  spread(movie, n, fill = 0) %>%
  gather(movie, count, 2:8) %>%
  mutate(total = sum(count)) %>%
  arrange(desc(total))
tarantino_swears
```

---

# 

```{r}
tarantino_swears %>%
  ungroup() %>%
  mutate(word = fct_reorder(word, total)) %>%
  hchart(type = "bar",
         hcaes(
           x = word,
           y = count,
           group = movie
         ))
```

---

# Highcharter customisation

The developer of highcharter has made the package as close as possible to the underlying JavaScript library.

The [api.highcharts.com](https://api.highcharts.com) pages inform us that to change the stacking of the bars we need to change the following options:

> plotOptions.series.stacking

In R we do the following:

```{r, eval=FALSE}
tarantino_swears %>%
  ungroup() %>%
  mutate(word = fct_reorder(word, total)) %>%
  hchart(type = "bar",
         hcaes(
           x = word,
           y = count,
           group = movie
         )) %>%
  hc_plotOptions(series = list(stacking = "normal"))
```

---

```{r, echo=FALSE}
tarantino_swears %>%
  ungroup() %>%
  mutate(word = fct_reorder(word, total)) %>%
  hchart(type = "bar",
         hcaes(
           x = word,
           y = count,
           group = movie
         )) %>%
  hc_plotOptions(series = list(stacking = "normal")) %>%
  hc_size(height = "100%")
```

---

# Exercise

Create a new project which achieves the following:

1. Reproducibly cleans the raw data from [martinjhnhadley/collated-rstats-training-materials/german-migrants](https://goo.gl/2LwAWh)

1. Import the cleaned data

1. Use either `plotly` or `highcharter` to visualise the breakdown of receive countries from each send country.

1. Barcharts are one method for displaying this data, but feel free to explore or suggest alternatives.
