---
title: "Interactive Charts with plotly"
author: "Martin John Hadley"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library("tidyverse")
library("fivethirtyeight")
library("plotly")
tarantino_swears <- tarantino %>%
  filter(complete.cases(.)) %>%
  count(word, movie) %>%
  group_by(word) %>%
  mutate(total.n = sum(n)) %>%
  filter(total.n > 1) 
```


# plotly

- `plot_ly` creates a plotly object to which we add layers
- `add_traces` automatically decides the series type based on the data
- columns must be specified as formula (~column.name)

```{r message=FALSE, warning=FALSE, eval = FALSE}
tarantino_swears %>%
  plot_ly() %>%
  add_trace(y = ,
            x = ,
            color = )
```

---

```{r message=FALSE, warning=FALSE, echo = FALSE}
tarantino_swears %>%
  plot_ly() %>%
  add_trace(y = ~word,
            x = ~n,
            color = ~movie)
```

---

# Ordering with factors

```{r, eval = FALSE}
tarantino_swears %>%
  plot_ly() %>%
  add_trace(y = ~fct_reorder(word, total.n),
            x = ~n,
            color = ~movie)
```


---

```{r, echo = FALSE}
tarantino_swears %>%
  plot_ly() %>%
  add_trace(y = ~fct_reorder(word, total.n),
            x = ~n,
            color = ~movie)
```

---

# layout

Unfortunately, plotly isn't that smart aboyt sizing axes labels. We use the `layout` function to affect how the chart appears:

```{r, eval = FALSE}
tarantino_swears %>%
  plot_ly() %>%
  add_trace(y = ~fct_reorder(word, total.n),
            x = ~n,
            color = ~movie) %>%
  layout(margins = list(l = 100))
```

---

```{r echo=FALSE, message=FALSE, warning=FALSE}
tarantino_swears %>%
  plot_ly() %>%
  add_trace(y = ~fct_reorder(word, total.n),
            x = ~n,
            color = ~movie) %>%
  layout(margins = list(l = 100))
```

---

# Stacking with barmode

```{r, eval = FALSE}
tarantino_swears %>%
  plot_ly() %>%
  add_trace(y = ~fct_reorder(word, total.n),
            x = ~n,
            color = ~movie) %>%
  layout(margins = list(l = 100),
         barmode = "stack")
```

---

## Learning more about plotly

We've not really scratched the surface of what's possible with `plotly`, there are a vast range of different visualisation options available. See [https://plot.ly/r/](https://plot.ly/r/) for comprehensive showcase of the different chart types you can use.

There's an online book about `plotly` currently being drafted here - [cpsievert.github.io/plotly_book/](http://cpsievert.github.io/plotly_book/). I recommend this with a bit of caution at the moment, it's not yet complete.

Of course, there are plenty of questions about the `plotly` library on StackOverflow: [https://stackoverflow.com/questions/tagged/plotly+r](https://stackoverflow.com/questions/tagged/plotly+r). 

Remember if you ask your own questions on StackOverflow to provide a [reproducible example](https://stackoverflow.com/a/5963610/1659890) of your problem.