---
title: "Interactive Networks with htmlwidgets"
author: "Martin John Hadley"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

# Interactive networks with htmlwidgets

There are a few different packages for creating interactive network diagrams, the only one we're going to cover is `visNetwork`.

Before doing so, it's important we cover some basic points about network charts.

---

# Network layouts are largely aesthetic

In general, the relative position of nodes within a network visualisation does not provide material information about the importance of that node.

The next three slides contain a network diagram of exactly the same data but with very different layouts.

---

```{r message=FALSE, warning=FALSE, echo=FALSE, opts.width = "100px"}
library("tidyverse")
library("igraph")
library("tidygraph")
library("visNetwork")
library("RColorBrewer")
library("widgetframe")

got_nodes <- read_csv("data/GoT_nodes.csv")
got_edges <- read_csv("data/GoT_edges.csv")

superculture_colours <- tibble(
  superculture = unique(got_nodes$superculture),
  color = brewer.pal(12, "Paired")
)

got_nodes <- got_nodes %>%
  left_join(superculture_colours) %>%
  mutate(group = superculture)

got_network <- tbl_graph(nodes = got_nodes, edges = got_edges)
```

```{r}
got_network %>%
  as.igraph() %>%
  visIgraph() %>%
  visOptions(highlightNearest = TRUE) %>%
  visLegend(ncol = 2)
```

---


```{r}
got_network %>%
  as.igraph() %>%
  visIgraph(layout = "layout_on_grid") %>%
  visOptions(highlightNearest = TRUE) %>%
  visLegend(ncol = 2)
```

---

```{r}
got_network %>%
  as.igraph() %>%
  visIgraph(layout = "layout_with_lgl") %>%
  visOptions(highlightNearest = TRUE) %>%
  visLegend(ncol = 2)
```

---

# Network layouts are largely aesthetic

Most layout algorithms use non-deterministic physics simulations to decide on the end position of nodes.

This is why the same network produced on two different computers (or at two different points in time) look different to one another.

```{r, eval = FALSE}
library(igraphdata)
data("karate")
karate %>%
  visIgraph()
```

---

# visNetwork

`visNetwork` is an extremely fussy library, it needs tibbles with specific column names and positions in order to function.

.pull-left[
```{r, eval=FALSE}
visn_nodes <- tibble(
  id = 1:5,
  label = letters[1:5]
)
```
]

.pull-right[
```{r, eval = FALSE}
visn_edges <- tibble(
  from = c(2, 1, 3, 4, 5),
  to = c(1, 4, 5, 1)
)
```
]

---

```{r, echo = FALSE}
visn_nodes <- tibble(
  id = 1:5,
  label = letters[1:5]
)
visn_edges <- tibble(
  from = c(2, 1, 3, 4, 5),
  to = c(1, 4, 5, 1, 2)
)
```


```{r}
visNetwork(visn_nodes,
           visn_edges)
```

---

# Exercise

1. Load the collated-rstats-training-materials/got-network project

1. Reproducibly wrangle the data into shape for visNetwork

1. Visualise the network
