---
title: "Plotly"
output:
  slidy_presentation:
    theme: default
css: "www/teachingnotes.css"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
	eval = FALSE,
	message = FALSE,
	warning = FALSE,
	include = TRUE
)
```

## Plotly

Plotly is an excellent tool for creating interactive charts, plots and other visualisations. 

The underlying JavaScript library is called plotly.js, it's also useful to know that it's developers provide a point-and-click service for building charts at [plot.ly](plot.ly).

## Plotly and ggplot2 (I)

Plotly's super power of converting ggplot2 visualisations into interactive visualisations is particularly powerful.

## Plotly and ggplot2 (II)

If you're unfamiliar with ggplot2, it's an extraudinarily powerful "grammar of graphics" for creating static charts.

It's very much dependent on correctly using factors. For a streamlined approach to working with factors I thoroughly recommend loading the tidyverse library `forcats`

```{r, eval=TRUE}
library("tidyverse")
library("forcats")
starwars %>%
  filter(!is.na(species)) %>%
  group_by(species) %>%
  summarise(mean.height = mean(height, na.rm = TRUE)) %>%
  ggplot(aes(
    x = fct_reorder(species, mean.height),
    y = mean.height
  )) + geom_col() +
  coord_flip() +
  xlab("Star Wars Species") + ylab("Height (cm)")
```

## Plotly and ggplot2 (III)

This static chart can be made interactive simply by providing it to `ggplotly`.

However, this functionaliry 

```{r, eval = TRUE}
library("tidyverse")
library("forcats")
library("plotly")
gg_starwars_height_chart <- starwars %>%
  filter(!is.na(species)) %>%
  group_by(species) %>%
  summarise(mean.height = mean(height, na.rm = TRUE)) %>%
  ggplot(aes(
    x = fct_reorder(species, mean.height),
    y = mean.height
  )) + geom_col() +
  coord_flip() +
  xlab("Star Wars Species") + ylab("Height (cm)")
ggplotly(gg_starwars_height_chart)
```

## Plotly and Gapminder (I)

The gapminder dataset is a popular socio-economic dataset for examples, there's a dedicated R package that makes it easy to work with the data:

```{r, eval=TRUE}
library("gapminder")
head(gapminder)
```

## Plotly and Gapminder (II)

Let's create this chart:

```{r, echo=FALSE, eval=TRUE}
library("tidyverse")
library("gapminder")
library("plotly")
continent_mean_pop_growth <- gapminder %>%
  group_by(continent, year) %>%
  summarise(mean.pop = mean(pop))

plotly_cont_mean_pop <- continent_mean_pop_growth %>%
  plot_ly() %>%
  add_trace(x = ~year,
            y = ~mean.pop,
            color = ~continent,
            type = "scatter",
            mode = "lines+markers") %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Mean population of countries in continent"),
         title = "Mean population of countries within each continent over time (source: Gapminder)")
plotly_cont_mean_pop
```


<br>
Steps required:

- Calculate mean continental population
- Create a plotly chart
- Add sugar

<br>
Additional information:

- plot_ly() instantiates a plotly object
- add_trace() is used to add traces, and access columns from the data argument using ~ (just like `leaflet`)

## Plotly and the journeys dataset (I)

Let's load up the plotly-and-the-journey.RProj project from RSS_handouts folder, there's a file here that processes our data ready to be visualised using a bar chart.

Let's create this chart together:

```{r, echo=FALSE, eval=TRUE}
library("tidyverse")
library("forcats")
library("plotly")

journeys <- read_csv("data/detailed-journeys.csv")

end_location_frequencies <- journeys %>%
  group_by(end.city, end.country) %>%
  count() %>%
  arrange(desc(n)) %>%
  ungroup() %>%
  mutate(end.city = fct_reorder(end.city, n))

my_plotly <- end_location_frequencies %>%
  plot_ly(width = "100%") %>%
  add_trace(
    y = ~end.city,
    x = ~n,
    color = ~end.country
  ) %>%
  layout(margin = list(l = 170),
         xaxis = list(title = "Number of observations"),
         yaxis = list(title = ""),
         title = "Receive city incidence in the journeys dataset")
my_plotly
```

## Learning more about leaflet

We've not really scratched the surface of what's possible with `plotly`, there are a vast range of different visualisation options available. See [https://plot.ly/r/](https://plot.ly/r/) for comprehensive showcase of the different chart types you can use.

There's an online book about `plotly` currently being drafted here - cpsievert.github.io/plotly_book/. I recommend this with a bit of caution at the moment, it's not yet complete.

Of course, there are plenty of questions about the `plotly` library on StackOverflow: [https://stackoverflow.com/questions/tagged/plotly+r](https://stackoverflow.com/questions/tagged/plotly+r). Remember if you ask your own questions on StackOverflow to provide a [reproducible example](https://stackoverflow.com/a/5963610/1659890) of your problem.